Procurement Dynamics: Analyzing Federal Contract Awards to Microsoft

### 2024-12-03 paper/paper.qmd
Update Data Section

More information about solicitation method can be found in @sec-appx.

@@

After imported the raw dataset using `read_csv` function, I first select specific columns to focus on relevant information towards this study by omitting redundant information, such as the contract link and same value of region. Then I rename key variables from raw data for clarity. The key variables of interest in this analysis include Contract, StartDate, AwardDate, EndDate, Buyer, and Amount.

For any missing data (NA) of StartDate, I infer the StartDate from the AwardDate to replace the missing value as the value of AwardDate. Then I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes. Due to the error with unknown reason that EndDate is before StartDate or AwardDate, those entries were excluded from this paper.

PreparatoryPhase, PhaseDays, and ContractDays variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract. PreparatoryPhase calculates the days from AwardDate to StartDate. ContractDays calculates the days from StartDate to EndDate. PhaseDays take the absolute value of PreparatoryPhase because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded. I also created new variables called ContractType and BuyerCleaned, which categorizes the most common buyers and contracts based on keywords.

This dataset contains only two entries for 2019 federal procurement contracts awarded to Microsoft. So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired. By summary the data set, I noticed a significant outlier with value of $102,260,150, which largely impact the outcome. So, I decide to drop this entry. Then I check the character type variables, Supplier and Contract, using `table()` function for the purpose of manipulating input error. I standardize the supplier's name which combine the similar name due to capitalization or some notation differences, and format them with first letter of each word capitalized.

@@

**Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract. [@tbl-buyer] shows the top 10 buyer's name who has over 50 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors. The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards. [CALCULATED 52.02%, BUT THINKING IF THIS SHOULD BE ADDED TO THE TABLE]

>>>>

More information about solicitation method can be found in @sec-meth-details.

@@

After imported the raw dataset using `read_csv` function, I first select specific columns to focus on relevant information towards this study by omitting redundant information, such as the contract link and same value of region. Then I rename key variables from raw data for clarity. The key variables of interest in this analysis include $Contract$, $StartDate$, $AwardDate$, $EndDate$, $Buyer$, and $Amount$.

I standardize the date format of all date related datas for proper parsing by using `mdy()` function and change the dollar format of amount into numeric numbers for future calculation purposes. Due to the error with unknown reason that $EndDate$ is before $StartDate$ or $AwardDate$, those entries were excluded from this paper.

$PreparatoryPhase$, $PhaseDays$, and $ContractDays$ variables were created for the purpose of analyzing preparatory phase and performance phase corresponding to each contract. $PreparatoryPhase$ calculates the days from $AwardDate$ to $StartDate$. $ContractDays$ calculates the days from $StartDate$ to $EndDate$. $PhaseDays$ take the absolute value of $PreparatoryPhase$ because of the minor effect of either prepare before contract being awarded or prepare after contract being awarded. I also created new variables called $ContractType$ and $BuyerCleaned$, which categorizes the most common buyers and contracts based on keywords.

This dataset contains only two entries for 2019 federal procurement contracts awarded to Microsoft. So, I filter out the contracts before 2020 to focus on the records between 2020 and September 2024 as desired. Also, I filtered missing data (NA) of $StartDate$, which only has one entry in this dataset. By summary the data set, I noticed a significant outlier with value of $102,260,150, which largely impact the outcome. So, I decide to drop this entry. Then I check the character type variables, $Supplier$ and $Contract$, using `table()` function for the purpose of manipulating input error. I standardize the supplier's name which combine the similar name due to capitalization or some notation differences, and format them with first letter of each word capitalized.

@@

**Buyer** is a character-type variable representing the federal government department or agency who initiate and award the contract. [@tbl-buyer] shows the top 10 buyer's name who has over 25 contracts with Microsoft, which identifies key partners and understand the distribution of contracts across sectors. The table highlights that Employment and Social Development Canada (ESDC) and National Defence account over 50% of Microsoft's total awards.
